Shape Enhanced With and Adapted Laplacian Operator For Hybrid Quad/Triangle Meshes

### 2013-01-15 documents/thesis/TQLBO-paper/Paper-Siggraph/paper.tex

\teaser{\includegraphics[width=1\textwidth]{images/spectrum} \caption{\label{fig:Spectrum}A set of 48 successive curvature enhance shapes,
from $\lambda=0.0$ in blue to $\lambda=-240.0$ in red, with steps
of $-5.0$.}
}

\maketitle
\maketitle 
\begin{abstract}
This paper proposes a novel modeling method for a hybrid quad/triangle
mesh that allows to set a family of possible shapes by controlling a single parameter, the global curvature. The method uses an
original extension of the Laplace Beltrami operator that efficiently estimates a curvature parameter which is used to define an enhanced shape after a particular operation performed in certain mesh points.  Along with the
method, this work presents new applications in sculpting and modeling,
with subdivision of surfaces and weight vertex groups. A series of
graphics examples demonstrates the quality, predictability and flexibility
of the method in a real production environment with software blender.
\end{abstract}

@@

Many tools have been developed for modeling, based on the Laplacian
mesh processing. Thanks to the advantages of the Laplacian operator,
these different tools preserve the surface geometric details when
using them for different processes such as free-form deformation,
fusion, morphing and other application\cite{Sorkine2004}.

@@

Gal et al. \cite{Gal2009} propose automatic feature detection and
shape edition with feature inter-relationship preservation. They define
salient surface features like ridges and valleys, characerized by
their first and second order curvature derivatives \cite{Ohtake2004},
and angle-based threshold. Likewise, curves have been also classified
as planar or non-planar, approximated by lines, circles, ellipses
and other complex shapes. In such case, the user defines an initial
change over several features which is propagated towards other features,
based on the classified shapes and the inter-relationships between
them. This method works well with objects that have sharp edges, composed
of basic geometric shapes such as lines, circles or ellipses. However,
the method is very limited when models are smooth since it cannot
find the proper features to edit.

@@

Many smoothing Laplacian functionals regularize the surface energy
by controling the total surface curvature$S$.

@@

Applying the mean average area, according to \cite{Xiong2011}, from
all possible triangulations, as show in figure \ref{fig:quad_xiong},
the area for each quad $\ensuremath{A\left(Q_{N_{1}\left(v_{i}\right)}\right)}$
is

@@

allows to move the new vertices to:

@@

\caption{\label{fig:Catmull_Clark}(a) Original Model, (b) Model with Catmull-Clark
Subdivision. Models with Laplacian smoothing: (c) $\lambda=60.0$,
$\lambda_{e}=12.0$ and (d) $\lambda=90.0$, $\lambda_{e}=80.0$.
Models with a first Laplacian filtering $\lambda=60.0$, $\lambda_{e}=12.0$
and before applying curvature enhancement: (e) $\lambda=-200.0$,
$\lambda_{e}=12.0$ and (f) $\lambda=-40.0$, $\lambda_{e}=-100.0$.}
\end{figure*}

A new sculpting brush is herein proposed and aims to inflate the shape,
magnifying the shape curvatures of a polygon mesh in real time. This
brush works properly with the stroke method \textsl{Drag Dot}, allowing
to pre-visualize the model changes before the mouse is released. Also,
it allows to move the mouse along the model to match the curvature
zone which is supposed to be enhanced.

@@

Laplacian smoothing applied with simple subdivision may produce similar
results to those obtained with Catmull-Clark, whose models are in
average equal triangles, as shown in figure \ref{fig:Catmull_Clark},
(b) Catmull-Clark, (c) Laplacian smoothing. The one obtained with
the Laplacian smoothing is shown in panel (c), (d) and those curvature
modified versions are in (e) and (f). As can be observed, different
versions of the original sketch can be obtained by parameterizing
a single model value, a great advantage of the presented method. Figure
\ref{fig:camello_enhanced} shows the generation of different versions
of a camel according to the $\lambda$ parameter. In the top row,
it is shown the curvature enhancement results, as $\lambda$ becomes
larger and negative, the resultant shape is observed as if the model
would inflate the more convex parts, as shown in figure \ref{fig:Spectrum}.
The larger the $\lambda$ parameter the larger the model feature enhancements.
The bottom row of figure \ref{fig:camello_enhanced} shows the use
of weighted vertex groups, specifying which areas will be enhanced.
On the left, the enhancement of the camel legs produces an organic
aspect, notice that the border is not distorted and smooth.

@@

Figure \ref{fig:Sculpt_Brush} shows the use of a curvature enhancement
brush for sculpting in real time. One pass was used with the brush,
as shown in the figure, with the blue and red radius. In figure \ref{fig:Sculpt_Brush}.b
the camel foot shows the inflation intersection that looks like two
bubbles, a similar pattern to what is observed to the fingers on the
bottom of the same figure. The silhouette enhancement is observed
in figure \ref{fig:Sculpt_Brush}.c since the main shape is retained
together with its finger and leg details. Similar results can be obtained
by a user, however it would take several steps and require the use
of several brushes, while the curvature enhancement took a single
step. Likewise, this new method can easily enhance organic features
like muscles during the sculpting process. In figure \ref{fig:Performance-sculpt}
the curvature enhancement brush performance is illustrated, in this
experiment three models with 12K, 40K and 164K vertices, were used.
These models were sculpted with the curvature enhance brush, at each
step the brush sphere containing a variable number of vertices for
processing. The processing time for 800 vertices in the camel paw
(40k model) only took 0.1 seconds, for 2600 vertices in the leg and
neck (model 40k) took 0.5 seconds, these times are suitable in real
applications since an artist sculpts a model for parts, and each part
is represented by an average of about 3000 vertices.

Tests with the Laplacian operator (equation \ref{eq:TQLBO_Simple_Matrix})
and its normalized version (equation \ref{eq:TQLBO-Normalized_Matrix}),
produce similar results if the triangles that compose the mesh are
about the same size. The normalized version is more stable and predictable
because it is not divided by the area of the ring which may be very
small and causes numeric problems, as shown in figure \ref{fig:(a)Monkey}.c
bottom row. The curvature enhancement of the model with the normalized
Laplacian operator results in a more regular pattern. The model can
be deformed with a TQLBO normalized version with large $\lambda$
($\lambda>400$) while intersecting itself with no peaks. Figure \ref{fig:(a)Monkey}
shows different results due to the triangle areas in the model. Triangles
with larger area have smaller enhancement (figure \ref{fig:(a)Monkey}.c
skull), and smaller triangles have larger enhancements (figure \ref{fig:(a)Monkey}.c
chin).

@@

\caption{\label{fig:(a)Monkey}(a) Top row: Original model scaled by 4. Bottom
row: Original Model (b) Top and bottom row: enhancing with Normalized-TQLBO
$\lambda=-50$ (c) Top and bottom row: enhancing with TQLBO $\lambda=-50$.}
\end{figure}

@@

To improve the performance, it was worked with the Blender mesh struct,
visiting each triangle or quad and storing its corresponding index
and the sum of the Laplacian weights of the ring in a list so that
only two visits were required for the list of mesh faces and two times
the edge list, if the mesh was not closed. This drastically reduced
calculations, enabling real-time processing. In the construction of
the Laplacian matrix, several index were locked at vertices having
face areas or edge lengths with null value that could cause spikes
and bad results.

>>>>

\teaser{\includegraphics[width=1\textwidth]{images/spectrum} \caption{\label{fig:Spectrum}A set of 48 successive shapes enhanced, from
$\lambda=0.0$ in blue to $\lambda=-240.0$ in red, with steps of
$-5.0$.}
}

\maketitle
\maketitle 
\begin{abstract}
This paper proposes a novel modeling method for a hybrid quad/triangle
mesh that allows to set a family of possible shapes by controlling
a single parameter, the global curvature. The method uses an original
extension of the Laplace Beltrami operator that efficiently estimates
a curvature parameter which is used to define an enhanced shape after
a particular operation performed in certain mesh points. Along with
the method, this work presents new applications in sculpting and modeling,
with subdivision of surfaces and weight vertex groups. A series of
graphics examples demonstrates the quality, predictability and flexibility
of the method in a real production environment with software Blender.
\end{abstract}

@@

Many tools have been developed for modeling, based on the Laplacian
mesh processing. Thanks to the advantages of the Laplacian operator,
these different tools preserve the surface geometric details when
using them for different processes such as free-form deformation,
fusion, morphing and other applications \cite{Sorkine2004}.

@@

\cite{Gal2009} propose automatic feature detection and shape edition
with feature inter-relationship preservation. They define salient
surface features like ridges and valleys, characerized by their first
and second order curvature derivatives \cite{Ohtake2004}, and angle-based
threshold. Likewise, curves have been also classified as planar or
non-planar, approximated by lines, circles, ellipses and other complex
shapes. In such case, the user defines an initial change over several
features which is propagated towards other features, based on the
classified shapes and the inter-relationships between them. This method
works well with objects that have sharp edges, composed of basic geometric
shapes such as lines, circles or ellipses. However, the method is
very limited when models are smooth since it cannot find the proper
features to edit.

@@

Many smoothing Laplacian functionals regularize the surface energy
by controling the total surface curvature $S$.

@@

Applying the mean average area, according to \cite{Xiong2011}, from
all possible triangulations, as show in figure \ref{fig:quad_xiong},
the area for quads $\ensuremath{A\left(Q_{v_{i}}\right)}$ and triangles
$A\left(T_{v_{i}}\right)$ is.

@@

allows to scale the vertices to:

@@

\caption{\label{fig:Catmull_Clark}(a) Original Model, (b) Model with Catmull-Clark
Subdivision. Models with Laplacian smoothing: (c) and (d). Models
with a first Laplacian filtering $\lambda=60.0$, $\lambda_{e}=12.0$
and before applying shape enhancement: (e) and (f).}
\end{figure*}

A new sculpting brush is herein proposed and aims to inflate the shape,
magnifying the shape curvatures of a polygon mesh in real time. This
brush works properly with the stroke method \textsl{Drag Dot}, allowing
to pre-visualize the model changes before the mouse is released. Also,
it allows to move the mouse along the model to match the shape zone
which is supposed to be enhanced.

@@

Laplacian smoothing applied with simple subdivision (see figure \ref{fig:Catmull_Clark}.b.)
may produce similar results to those obtained with Catmull-Clark (see
figure \ref{fig:Catmull_Clark}.c.), whose models are in average equal
triangles. The one obtained with the Laplacian smoothing is shown
in panel (c), (d) and those curvature modified versions are in (e)
and (f). As can be observed, different versions of the original sketch
can be obtained by parameterizing a single model value, a great advantage
of the presented method. Figure \ref{fig:camello_enhanced} shows
the generation of different versions of a camel according to the $\lambda$
parameter. In the top row, it is shown the shape enhancement results,
as $\lambda$ becomes larger and negative, the resultant shape is
observed as if the model would inflate the more convex parts, as shown
in figure \ref{fig:Spectrum}. The larger the $\lambda$ parameter
the larger the model feature enhancements. The bottom row of figure
\ref{fig:camello_enhanced} shows the use of weighted vertex groups,
specifying which areas will be enhanced. On the left, the enhancement
of the camel legs produces an organic aspect, notice that the border
is not distorted and smooth.

@@

Figure \ref{fig:Sculpt_Brush} shows the use of a shape enhancement
brush for sculpting in real time. One pass was used with the brush,
as shown in the figure, with the blue and red radius. In figure \ref{fig:Sculpt_Brush}.b
the camel foot shows the inflation intersection that looks like two
bubbles, a similar pattern to what is observed to the fingers on the
bottom of the same figure. The silhouette enhancement is observed
in figure \ref{fig:Sculpt_Brush}.c since the main shape is retained
together with its finger and foot details. Similar results can be
obtained by a user, however it would take several steps and require
the use of several brushes, while the shape enhancement took a single
step. Likewise, this new method can easily enhance organic features
like muscles during the sculpting process. In figure \ref{fig:Performance-sculpt}
the shape enhancement brush performance is illustrated, in this experiment
three models with 12K, 40K and 164K vertices, were used. These models
were sculpted with the shape enhance brush, at each step the brush
sphere containing a variable number of vertices for processing. The
processing time for 800 vertices in the camel paw (40k model) only
took 0.1 seconds, for 2600 vertices in the leg and neck (model 40k)
took 0.5 seconds, these times are suitable in real applications since
an artist sculpts a model for parts, and each part is represented
by an average of about 1800 vertices.

Tests with the Laplacian operator (equation \ref{eq:TQLBO_Simple_Matrix})
and its normalized version (equation \ref{eq:TQLBO-Normalized_Matrix}),
produce similar results if the triangles or quads that compose the
mesh are about the same size. The normalized version is more stable
and predictable because it is not divided by the area of the ring
which may be very small and causes numeric problems, as shown in figure
\ref{fig:(a)Monkey}.c bottom row. The shape enhancement of the model
with the normalized Laplacian operator results in a more regular pattern.
The model can be deformed with a TQLBO normalized version with large
$\lambda$ ($\lambda>400$) while intersecting itself with no peaks.
Figure \ref{fig:(a)Monkey}.c shows different results due to the quads
areas in the model. Quadss with larger area have smaller enhancement
(figure \ref{fig:(a)Monkey}.c skull), and smaller quads have larger
enhancements (figure \ref{fig:(a)Monkey}.c chin).

@@

\caption{\label{fig:(a)Monkey}(a) Bottom row: Original Model. Top row: Original
model scaled by 4. (b) Top and bottom row: enhancing with Normalized-TQLBO
$\lambda=-50$ (c) Top and bottom row: enhancing with TQLBO $\lambda=-50$.}
\end{figure}

@@

To improve the performance, it was worked with the Blender mesh struct,
visiting each triangle or quad and storing its corresponding index
and the sum of the Laplacian weights of the ring in a list so that
only two visits were required for the list of mesh faces and two times
the edge list, if the mesh was not closed. This drastically reduced
calculations, enabling real-time processing. In the construction of
the Laplacian matrix, several index were locked at vertices having
face areas or edge lengths with zero value that could cause spikes
and bad results.
